Determining the alternatives for scalar implicature

### 2016-02-01 Paper/Paper.tex
reorg finished

How much of what we mean comes from the words that go unsaid? As
listeners, our ability to make precise inferences about a speaker's
intended meaning in context is indispensable to successful
communication. For example, listeners commonly enrich the meaning of the
scalar item \emph{some} to \emph{some but not all} in sentences like
``Alice ate some of the cookies'' (Grice, 1975; Horn, 1984; Levinson,
2000). These inferences, called \emph{scalar implicatures}, have been an
important test case for understanding pragmatic inferences more
generally. A Gricean account of this phenomenon assumes listeners reason
about a speaker's intended meaning by incorporating knowledge about A)
alternative scalar items a speaker could have used (such as \emph{all})
and B) the relative informativity of using such alternatives (Grice,
1975). According to this account, a listener will infer that the speaker
must have intended that Alice did not eat all the cookies because it
would have been underinformative to use the descriptor \emph{some} when
the alternative \emph{all} could have been used just as easily.

@@

To instantiate our signaling game with a tractable message set \(M\), in
our studies we adopt the world of restaurant reviews as our
communication game. We assume that speakers and listeners are trying to
communicate the number of stars in an online restaurant review (where
\(m \in \{1...5\}\)). We then use experiments to measure three
components of the model. First, to measure literal semantics
\({p_{L_0} (m \mid u)}\) we ask experiment participants to judge whether
a message is compatible with a particular meaning (Experiment 1).
Second, to generate a set of plausible alternative messages in \(U\), we
elicit alternatives directly (Experiment 2). Lastly, to obtain human
\(L_1\) pragmatic judgments, we ask participants to interpret a
speaker's utterances (Experiment 3).

\section{Experiment 1: Literal listener
task}\label{experiment-1-literal-listener-task}

@@

\caption[(Left) A trial from Experiment 1 (literal listener) with the target scalar "liked]{(Left) A trial from Experiment 1 (literal listener) with the target scalar "liked."(Right) A trial from Experiment 3 (pragmatic listener) with the target scalar "liked."}\label{fig:stimuli_exp1}
\end{figure*}
\end{CodeChunk}

Experiment 1 was conducted to approximate literal listener
semantics---\(p_{L_0}(m \mid u)\) in Equation (2)---for five pairs of
scalar items taken from {van Tiel} (2014). We include three conditions
in Experiment 1, corresponding to the sets of alternatives within a
scale that participants were presented with: two-alternatives
(``entailment''), four-alternatives, and five-alternatives. The
two-alternatives condition makes a test of the hypothesis that the two
members of the classic Horn (entailment) scale (Horn, 1972) are the only
alternatives necessary to predict the strength of listeners' pragmatic
inference. The four and five-alternatives conditions then add
successively more alternatives to test whether including a larger number
of alternatives will increase model
fit.\footnote{Note that alternatives in the four and five-alternatives conditions were chosen on the basis of Experiment 2, which was run chronologically after the two-alternative condition; all literal listener experiments are grouped together for simplicity in reporting.}
A secondary goal of Experiment 1 is to test whether the set of
alternatives queried during literal semantic elicitation impacts literal
semantic judgments. If it does we should see differences in these
judgments for shared items between experiments.

@@

The two-alternatives condition included only the two terms in the pairs
from van Tiel (2014). The four-alternatives condition included the two
scalar items plus the top two alternatives generated for each scalar
family by participants in Experiment 2. The five-alternatives condition
included the four previous items plus one more neutral item chosen from
alternatives generated in Experiment 2. See Figure
\ref{fig:allScalesTable} for the complete list of alternatives used in
each condition.

@@

\caption[Literal listener judgments from Experiment 1]{Literal listener judgments from Experiment 1. Proportion of participants indicating compatibility (answering "yes") is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Rows are grouped by scale and items within rows are ordered by valence. Colors indicate the specific condition (1a,b,c) with conditions including different numbers of items.}\label{fig:exp1Plots}
\end{figure*}
\end{CodeChunk}

\section{Experiment 2: Alternative
Elicitation}\label{experiment-2-alternative-elicitation}

Participants were presented a target scalar item from our original
entailment set (see Figure \ref{fig:allScalesTable}) embedded in a
sentence such as, ``In a recent restaurant review someone said they
thought they the food was \_\_\_\_," with a target scalar presented in
the blank. Participants were then asked to generate plausible
alternatives by responding to the question, ``If they'd felt differently
about the food, what other words could they have used instead of
\_\_\_\_?" They were prompted to generate three unique alternatives.

@@

\caption[Pragmatic listener judgements for scalar items]{Pragmatic listener judgements for scalar items. Proportion of participants generating a star rating is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Line type denotes condition, and colors indicate the particular scalar items. Each panel shows one scalar pair, with only entailment items (two-alternatives condition) shown here for simplicity in reporting.}\label{fig:exp2Plots}
\end{figure*}
\end{CodeChunk}

Figure \ref{fig:exp2_altsPlot_likedLoved} shows an example alternative
set for the scalar items \emph{liked} and \emph{loved} (combined). We
expanded our alternative set from only entailment items (items used in
Experiments 1a, 3a) to four alternatives (Experiments 1b, 3b) by
including the top two alternatives elicited for each scale in Experiment
2. For example, the top two alternatives for the
\emph{liked}/\emph{loved} scale were \emph{hated} and \emph{disliked}
and were included in four-alternatives model. Similarly, neutral items
(e.g. \emph{okay} in the \emph{good}/\emph{excellent} scale) were chosen
from among the set of alternatives elicited in Experiment 2 and added to
the five-alternatives model. See Figure \ref{fig:allScalesTable} for the
complete list of alternatives we used from Experiment 2.

@@

Figure \ref{fig:exp2Plots} plots pragmatic listener judgment
distributions for ``weak'' / ``strong'' scalar pairs (e.g.
\emph{good}/\emph{excellent}). Several trends are visible. First, in
each scale participants generated implicatures. They were substantially
less likely to assign high star ratings to weaker scalar terms, despite
the literal semantic compatibility of those terms with those states
shown in Experiment 1. Second, the size of the difference between strong
and weak scalar items varied considerably across scales, consistent with
previous work ({van Tiel}, 2014).

To rule out the potential effects of having a larger set of alternatives
during the pragmatic judgment elicitation, we fit a mixed effects model.
We regressed pragmatic judgments on scale and experimental condition
with subject- and word-level random effects, which was the maximal
structure that converged. There were no significant differences between
conditions 3a and 3b, \(b =\) 0.05, \(t(\) 150 \() =\) 1.04, \(p =\) 0.3
and the addition of the condition predictor did not significantly
improve model fit when compared to a model without that variable
\(\chi^2(\) 1 \() =\) 1.13, \(p =\) 0.29.

@@

Using literal listener data from Experiment 1, we conducted a set of
simulations with the RSA model. Each simulation kept the model constant,
fitting the choice parameter \(\alpha\) as a free parameter, but used a
set of alternatives to specify the scale over which predictions were
computed. We considered four different alternative sets, with empirical
measurements corresponding to those shown in Figure
\ref{fig:allScalesTable}: 1) the two-alternatives in the classic
entailment scales, 2) those two alternatives with the addition of a
generic negative
alternative\footnote{Literal semantics for the generic negative alternative were not generated empirically, but serve as a baseline ``none" style semantics in which the scalar item is only compatible with 1 star.},
3) the expanded set of four alternatives tested in Experiment 1b, and 4)
the expanded set of five alternatives tested in Experiment 1c.

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \hline
Model & $\alpha$ & Exp. 3a & Exp. 3b \\ 
  \hline
Two alts &   9 & 0.54 & 0.57 \\ 
  Two alts + generic negative  &   6 & 0.62 & 0.66 \\ 
  Four alts &   4 & 0.84 & 0.90 \\ 
  Five alts &   4 & 0.86 & 0.90 \\ 
   \hline
\end{tabular}
\caption{Model performance with fitted alpha levels. Model fit assessed through correlation with human judgments in our two Pragmatic listener experiments (3a,b)} 
\end{table}

@@

In our current work, we used the rational speech act framework to
investigate the set of alternatives that best allowed the model to
predict pragmatic judgments across a series of different scales. We
found that the best predictions came when a range of scale-dependent
negative and neutral alternatives were added to the implicature
computation, suggesting the importance of considering non-entailment
alternatives. This work builds on previous investigations, reinforcing
the claim that negative alternatives are critical for understanding
implicature (Franke, 2014), and replicates and extends findings that
different lexical scales produce strikingly different patterns of
inference ({van Tiel}, 2014).

While improvements in model fit were substantial as we moved from two to
four alternatives, we saw only a minor increase in fit from the move to
five alternatives. One possible explanation is that alternatives are
differentially salient in context, and in moving to larger sets we
should consider weighting the alternatives differentially (as Franke,
2014 did). Preliminary simulations using weightings derived from
Experiment 2 provide some support for this idea but would require
further empirical work for confirmation.

@@

Horn, L. R. (1984). Toward a new taxonomy for pragmatic inference:
Q-based and R-based implicature. \emph{Meaning, Form, and Use in
Context}, \emph{42}.

>>>>

How much of what we mean comes from the words that go unsaid? As
listeners, our ability to make precise inferences about a speaker's
intended meaning in context is indispensable to successful
communication. For example, listeners commonly enrich the meaning of the
scalar item \emph{some} to \emph{some but not all} in sentences like
``Alice ate some of the cookies'' (Grice, 1975; Levinson, 2000). These
inferences, called \emph{scalar implicatures}, have been an important
test case for understanding pragmatic inferences more generally. A
Gricean account of this phenomenon assumes listeners reason about a
speaker's intended meaning by incorporating knowledge about A)
alternative scalar items a speaker could have used (such as \emph{all})
and B) the relative informativity of using such alternatives (Grice,
1975). According to this account, a listener will infer that the speaker
must have intended that Alice did not eat all the cookies because it
would have been underinformative to use the descriptor \emph{some} when
the alternative \emph{all} could have been used just as easily.

@@

To instantiate our signaling game with a tractable message set \(M\), in
our studies we adopt the world of restaurant reviews as our
communication game. We assume that speakers and listeners are trying to
communicate the number of stars in an online restaurant review (where
\(m \in \{1...5\}\)). We then use experiments to measure three
components of the model. First, to generate a set of plausible
alternative messages in \(U\), we ask paticipants to generate
alternatives for particular scalar items (Experiment 1). Next, to
measure literal semantics \({p_{L_0} (m \mid u)}\) we ask experiment
participants to judge whether a message is compatible with a particular
meaning (Experiment 2). Lastly, to obtain human \(L_1\) pragmatic
judgments, we ask participants to interpret a speaker's utterances
(Experiment 3).

@@

\section{Experiment 1: Alternative
Elicitation}\label{experiment-1-alternative-elicitation}

@@

Participants were presented with a target scalar item from five scales,
selected from {van Tiel} et al. (2014) (see Figure
\ref{fig:allScalesTable}). These were embedded in a sentence such as,
``In a recent restaurant review someone said they thought they the food
was \_\_\_\_," with a target scalar presented in the blank. Participants
were then asked to generate plausible alternatives by responding to the
question, ``If they'd felt differently about the food, what other words
could they have used instead of \_\_\_\_?" They were prompted to
generate three unique alternatives.

@@

@@

\caption[(Left) A trial from Experiment 2 (literal listener) with the target scalar "liked]{(Left) A trial from Experiment 2 (literal listener) with the target scalar "liked."(Right) A trial from Experiment 3 (pragmatic listener) with the target scalar "liked."}\label{fig:stimuli_exp1}
\end{figure*}
\end{CodeChunk}

Figure \ref{fig:exp2_altsPlot_likedLoved} shows an example alternative
set for the scalar items \emph{liked} and \emph{loved} (combined).
Figure \ref{fig:allScalesTable} shows the complete list of alternative
sets derived from Experiment 1.

\section{Experiment 2: Literal listener
task}\label{experiment-2-literal-listener-task}

Experiment 2 was conducted to approximate literal listener
semantics---\(p_{L_0}(m \mid u)\) in Equation (2)---for the same five
pairs of scalar items used in Experiment 1. We included three
conditions: two alternatives (``entailment''), four alternatives, and
five alternatives. The two alternatives condition makes a test of the
hypothesis that the two members of the classic Horn (entailment) scale
(Horn, 1972) are the only alternatives necessary to predict the strength
of listeners' pragmatic inference. The four and five alternatives
conditions then add successively more alternatives to test whether
including a larger number of alternatives will increase model
fit.\footnote{Note that alternatives in the four and five alternatives conditions were chosen on the basis of Experiment 1, which was run chronologically after the two-alternative condition; all literal listener experiments are grouped together for simplicity in reporting.}
A secondary goal of Experiment 2 is to test whether the set of
alternatives queried during literal semantic elicitation impacts literal
semantic judgments. If it does we should see differences in these
judgments for shared items between experiments.

The two-alternatives condition included only the two terms in the pairs
from van Tiel (2014). The four-alternatives condition included the two
scalar items plus the top two alternatives generated for each scalar
family by participants in Experiment 1. The five-alternatives condition
included the four previous items plus one more neutral item chosen from
alternatives generated in Experiment 1.

\caption[Literal listener judgments from Experiment 2]{Literal listener judgments from Experiment 2. Proportion of participants indicating compatibility (answering "yes") is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Rows are grouped by scale and items within rows are ordered by valence. Colors indicate the specific condition (1a,b,c) with conditions including different numbers of items.}\label{fig:exp1Plots}
\end{figure*}
\end{CodeChunk}

@@

\caption[Pragmatic listener judgements in Experiment 3]{Pragmatic listener judgements in Experiment 3. Proportion of participants generating a star rating is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Line type denotes condition, and colors indicate the particular scalar items. Each panel shows one scalar pair, with only entailment items (two alternatives condition) shown here for simplicity in reporting.}\label{fig:exp2Plots}
\end{figure*}
\end{CodeChunk}

Figure \ref{fig:exp2Plots} plots pragmatic listener judgment
distributions for ``weak'' / ``strong'' scalar pairs (e.g.
\emph{good}/\emph{excellent}). Several trends are visible. First, in
each scale participants generated implicatures. They were substantially
less likely to assign high star ratings to weaker scalar terms, despite
the literal semantic compatibility of those terms with those states
shown in Experiment 2. Second, the size of the difference between strong
and weak scalar items varied considerably across scales, consistent with
previous work ({van Tiel} et al., 2014).

To rule out the potential effects of having a larger set of alternatives
during the pragmatic judgment elicitation, we fit a mixed effects model.
We regressed pragmatic judgments on scale and experimental condition
with subject- and word-level random effects, which was the maximal
structure that converged. There were no significant differences between
the two alternatives and four alternatives conditions (\(\beta = 0.05\),
\(t(150) = 1.04\), \(p = 0.3\)). The addition of the condition predictor
did not significantly improve model fit when compared to a model without
that variable (\(\chi^2(1) = 1.13\), \(p = 0.29\)).

@@

Using literal listener data from Experiment 2, we conducted a set of
simulations with the RSA model. Each simulation kept the model constant,
fitting the choice parameter \(\alpha\) as a free parameter, but used a
set of alternatives to specify the scale over which predictions were
computed. We considered four different alternative sets, with empirical
measurements corresponding to those shown in Figure
\ref{fig:allScalesTable}: 1) the two-alternatives in the classic
entailment scales, 2) those two alternatives with the addition of a
generic negative alternative, 3) the expanded set of four alternatives,
and 4) the expanded set of five alternatives. Literal semantics for the
generic negative alternative serve as a baseline ``none``-style
semantics in which the scalar item is only compatible with 1 star.

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \hline
Model & $\alpha$ & Two alts & Four alts \\ 
  \hline
Two alts &   9 & 0.54 & 0.57 \\ 
  Two alts + generic negative  &   6 & 0.62 & 0.66 \\ 
  Four alts &   4 & 0.84 & 0.90 \\ 
  Five alts &   4 & 0.86 & 0.90 \\ 
   \hline
\end{tabular}
\caption{Model performance with fitted alpha levels. Model fit assessed through correlation with human judgments in the two conditions of Experiment 3.} 
\end{table}

@@

In our current work, we used the rational speech act framework to
investigate the set of alternatives that best allowed the model to
predict pragmatic judgments across a series of different scales. We
found that the best predictions came when a range of scale-dependent
negative and neutral alternatives were added to the implicature
computation, suggesting the importance of considering non-entailment
alternatives. This work builds on previous investigations, reinforcing
the claim that negative alternatives are critical for understanding
implicature (Franke, 2014), and replicates and extends findings that
different lexical scales produce strikingly different patterns of
inference ({van Tiel} et al., 2014).

While improvements in model fit were substantial as we moved from two to
four alternatives, we saw only a minor increase in fit from the move to
five alternatives. One possible explanation is that alternatives are
differentially salient in context, and in moving to larger sets we
should consider weighting the alternatives differentially (as Franke,
2014 did). Preliminary simulations using weightings derived from
Experiment 1 provide some support for this idea but would require
further empirical work for confirmation.

@@
